Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-11-09 edubots_surveys.Rnw
Colouring changes
I forgot to do it before. :_D

Our intention from the beginning was to try to gather a sufficient amount of
responses in all sectors, as well as cover different sectors of higher
education. In doing so, an online survey was administered (i.e., Google Form),
in Spanish, which was intended to be answered in a very short amount of
time. Initially, our intention was to mirror the questions used in
\cite{MoraChatbots2021:anon}; however, an initial exploration of results led us
to create a second survey. The survey items focused on the use of messaging apps
in the teaching practice, the type of chatbot use cases that teachers consider
useful to be used in their teaching, and the type of interactions that teachers
prefer to have with their students while using messaging apps in their teaching.

@@

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). Survey
questions are shown in Appendix I (Section \ref{app:questions}).

@@

Group messaging apps have become a powerful tool for \textcolor{blue}{communicating} with
students. Also, many of them have built-in chatbots to enhance the learning
process. After an initial analysis to the answers to the first survey, we
realized that there were some prior issues mainly related to the adoption of a
technology, i.e. chatbots, that generally piggybacks on another, i.e. messaging
applications. In a technology adoption model, perceived ease of use and
perceived usefulness are essential. But it is going to be very difficult for
teachers to find {\em chatbots} easy to use if they, previously, do not
consider, or simply use, the messaging tools to which they are attached.

This is why for answering RQ1, teachers were queried about whether they use
messaging apps in their classes, specifically, Telegram, WhatsApp, Slack (an
application used mainly in IT departments and software development), or any
other messaging app, and whether they use messaging apps provided by their
academic institution (see Table \ref{tab:use_messapps}). In this case, we would like 
to add a clarification. While we had in mind, as indicated by the possible answers, 
instant messaging applications when making this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em source} of the technology (university, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based systems integrated in another learning management system).

@@

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use},
\ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not
show significant differences in the use of instant messaging apps between
teachers from University and vocational education. In general, most teachers
prefer messaging apps provided by their own institutions (we also show the
results for the two most popular messaging app platforms: WhatsApp and
Telegram). With respect to specific disciplines, Engineering and Technology
teachers are more active in their use, but the amount of teachers from Humanities 
who answered they used these apps in their classes is also remarkable
(around 60\% use the apps provided by their institutions).

>>>>

Our intention from the beginning was to try to gather a sufficient amount of
responses in all sectors, as well as cover different sectors of higher
education. In doing so, an online survey was administered (i.e., Google Form),
in Spanish, which was intended to be answered in a very short amount of
time. Initially, our intention was to mirror the questions used in
\cite{MoraChatbots2021:anon}; however, an initial exploration of results led us
to create a second survey. The \textcolor{blue}{new survey} items focused on the use of messaging apps
in the teaching practice, the type of chatbot use cases that teachers consider
useful to be used in their teaching, and the type of interactions that teachers
prefer to have with their students while using messaging apps in their teaching.

@@

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). Survey
questions \textcolor{blue}{are shown in Appendix I (Section \ref{app:questions})}.

@@

Group messaging apps have become a powerful tool for \textcolor{blue}{communicating} with
students. Also, many of them have built-in chatbots to enhance the learning
process. After an initial analysis to the answers to the first survey, we
realized that there were some prior issues mainly related to the adoption of a
technology, \textcolor{blue}{i.e. chatbots}, that generally piggybacks on another, \textcolor{blue}{i.e. messaging
applications}. In a technology adoption model, perceived ease of use and
perceived usefulness are essential. But it is going to be very difficult for
teachers to find {\em chatbots} easy to use if they, previously, do not
consider, or simply use, the messaging tools to which they are attached.

This is why for answering RQ1, teachers were queried about whether they use
messaging apps in their classes, specifically, Telegram, WhatsApp, Slack (an
application used mainly in IT departments and software development), or any
other messaging app, and whether they use messaging apps provided by their
academic institution (see Table \ref{tab:use_messapps}). \textcolor{blue}{In this case, we would like 
to add a clarification}. While we had in mind, as indicated by the possible answers, 
instant messaging applications when making this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em source} of the technology (university, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based systems integrated in another learning management system).

@@

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use},
\ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not
show significant differences in the use of instant messaging apps between
teachers from University and vocational education. In general, most teachers
prefer messaging apps provided by their own institutions (we also show the
results for the two most popular messaging app platforms: WhatsApp and
Telegram). With respect to specific disciplines, Engineering and Technology
teachers are more active in their use, \textcolor{blue}{but the amount of teachers from Humanities 
who answered they used these apps in their classes is also remarkable}
(around 60\% use the apps provided by their institutions).
